Parameter-Free Exploration in Deep Reinforcement Learning

### 2022-01-21 papers/main.tex
TEMPORARY DEPLOY COMMIT

%% Some suggested packages, as needed:
\usepackage{natbib} % has a nice set of citation styles and commands
    \bibliographystyle{plainnat}
    \renewcommand{\bibsection}{\subsubsection*{References}}
\usepackage{mathtools} % amsmath with fixes and additions
% \usepackage{siunitx} % for proper typesetting of numbers and units
\usepackage{booktabs} % commands to create good-looking tables
\usepackage{tikz} % nice language for creating drawings and diagrams
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{bm}

@@

\title{Scalable and Parameter-Free Exploration in Deep Reinforcement Learning}

@@

The efficiency of exploration can be measured in regret.
The regret of a policy is the difference in the expected reward obtained by
following that policy, and the expected reward of following an optimal policy.
An learning algorithm's efficiency in exploration can be measured by its
cumulative regret over time. One of the simplest reinforcement learning 
problems is the multi-arm bandit problem. This is an MDP with no state and no
transition probabilities. One exploration strategy employed in most Q-learning
algorithms to date, \(\epsilon\)-greedy is provably inefficient, and has a
regret bound that grows linearly with time. There are several optimal algorithms
for this problem, but perhaps the simplest one is Thompson sampling~\citep{thompson_likelihood_1933}.
Thompson sampling approximates a posterior distribution of the expected reward
for each action. The next action is decided by sampling rewards for each action
from the posterior distribution, and selecting the action that gave the highest
reward. Bayesian methods been also been shown to behave efficiently with respect
to cumulative regret \cite{osband_generalization_2016} on general MDPs.
%\textcolor{red}{Optimal Regret + Thompson sampling }

The \(Q\)-function, \(Q^\pi(s,a)\), is defined as the expected reward of taking action
\(a\) in state \(s\) and then following policy \(\pi\) thereafter:
\(Q^\pi \defeq \E_\pi \left[ \sum_{t=0}^\infty \gamma^t r_t \vert s_0=S, a_0=A \right]\).
The Bellman operator on \(Q^\pi\) is defined as
\(\B[Q^\pi](s_t,a_t) \defeq \E_{P(s_{t+1} \vert s_t, a_t)\pi(a_{t+1} \vert s_t)}
\left[ r(s_t, a) + \gamma Q^\pi(s_{t+1}, a_{t+1}) \right]\). With the
\(Q\)-function we can define a policy that always picks the action with the 
highest \(Q\)-value in any state. In deep RL, we model \(Q\) with a
neural network \(\hat{Q}_\omega\). One way to facilitate exploration in this
policy is to introduce uncertainty in the
\(Q\)-value function.

@@

We would like to approximate the posterior distribution of the Q-value function.
\citet{osband_why_2017} proove that in finite horizon MDPs posterior sampling for
reinforcement learning has at least a near-optimal regret bound. They further conjecture
that this cound can be further improved to show an optimal regret bound. Additionally,
a posterior sampling based reinfrocement learning algorithm could be made to utilize
prior information or domain knowledge through an appropriate prior. This should improve
policy convergence rate.
Combined with the computational efficiency of posterior sampling, this motivates the
development of a Bayesian reinforcement learning algorithm with functional priors.

We can use the functional variational bayesian neural
network (FVBNN)~\citep{sun_functional_2019} framework discussed earlier.
NoisyNet uses one sample from Q for each optimization step. We will instead use
\(N\) samples from Q. This lets us add the KL-term from FVBNN to the loss function that
encourages stochasticity.
The maximization target becomes:
\begin{equation}
    \log p(\D_\omega^N \vert f) - \KL \left[ q(\f^{\D_\omega^N}, \f^\mathbf{M}) \Vert p(\f^{\D_\omega^N}, \f^\mathbf{M}) \right],
\end{equation}
where \(\f^{\D_\omega^N}\) is \(f\) applied to the dataset \(\D_\omega^N\), and \(\f^\mathbf{M}\) is \(f\) applied to
a set of i.i.d. random points \(\mathbf{M} = \{m_i \sim c \mid i=1,\dots,k\} \subseteq \state \times \action\)
with full support. We will let the measurement \(\mathbf{M}\) be the the set of all state-action
pairs for states we have already explored. If the MDP is ergodic, then \(\textbf{M}\) will eventually
have full support in \(\state \times \action\).

@@

\begin{algorithm}
\caption{Functional Bayesian RL}
\begin{algorithmic}
    \State \(\D \gets \emptyset\)
    \State \(s \sim P_0\)
    \State Initialize \(\phi, \omega\)
    \While{not converged}
        \State \(f \sim q_\phi\)
        \State \(a \sim \arg\max_a f(s,a)\)
        \State \(s' \sim P(\cdot \vert s,a)\)
        \State \(r = r(s',a,s)\)
        \State \(\D \gets \D \cup \{s,a,r,s'\}\)
        \State \Call{UpdatePosterior}{$\phi, \omega, \D$}
    \EndWhile
    \Statex
    \Function{UpdatePosterior}{$\phi, \omega, \D$}
        \State \(T \sim \D\)
        \State \(f \gets \{f_i \sim q_\phi \mid i=1,\dots,k\}\)
        \State \(\mathbf{b} \gets \{b_i \sim q_\omega \mid i=1,\dots,k\}\)
        \State \(\mathcal{F} \gets \emptyset\)
        \State \(\Delta_\mathcal{L} \gets 0\)
        \For{\(\{s,a,r,s'\} \in T\)}
            \State \(\bm{q} \gets \max_a \f(s,a)\)
            \State \(\mathcal{F} \gets \mathcal{F} \cup \f(s,a)\)
            \State \(\bm{G} \gets \{r + \gamma b_i(s,a) \mid i = 1, 
        \dots, k\}\)
            \State \(\Delta_\mathcal{L} \gets \Delta_\mathcal{L} - \frac{1}{\lvert T \rvert} \nabla_\phi \log p(\bm{q} \vert \bm{G})\)
        \EndFor
        \State \(\Delta_\KL \gets SSGE(p, \mathcal{F})\)
        \State \(\phi \gets \phi + \eta_\phi (\Delta_\mathcal{L} - \lambda \Delta_\KL)\)
        \State \(\omega \gets \omega + \eta_\omega (\omega - \phi)\)
    \EndFunction
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{Example_Figure_1}
    \caption{Example figure 1}
    \label{fig:fig_1}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{example-image}
    \caption{Performance on the CartPole-v1 environment.}
\end{figure}

>>>>

%% Some suggested packages, as needed:
\usepackage{natbib} % has a nice set of citation styles and commands
    \bibliographystyle{plainnat}
    \renewcommand{\bibsection}{\subsubsection*{References}}
\usepackage{mathtools} % amsmath with fixes and additions
% \usepackage{siunitx} % for proper typesetting of numbers and units
\usepackage{booktabs} % commands to create good-looking tables
\usepackage{tikz} % nice language for creating drawings and diagrams
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{bm}
\usepackage{subcaption}

@@

\title{Parameter-Free Exploration in Deep Reinforcement Learning}

@@

The efficiency of exploration can be measured in regret.
The regret of a policy is the difference in the expected reward obtained by
following that policy, and the expected reward of following an optimal policy.
An learning algorithm's efficiency in exploration can be measured by its
cumulative regret over time. One of the simplest reinforcement learning 
problems is the multi-arm bandit problem. This is an MDP with no state and no
transition probabilities. One exploration strategy employed in most Q-learning
algorithms to date, \(\epsilon\)-greedy is provably inefficient, and has a
regret bound that grows linearly with time. There are several optimal algorithms
for this problem, but perhaps the simplest one is Thompson sampling~\citep{thompson_likelihood_1933}.
Thompson sampling approximates a posterior distribution of the expected reward
for each action. The next action is decided by sampling rewards for each action
from the posterior distribution, and selecting the action that gave the highest
reward. Bayesian methods has also been shown to behave efficiently with respect
to cumulative regret \cite{osband_generalization_2016} on general MDPs.
%\textcolor{red}{Optimal Regret + Thompson sampling }

The \(Q\)-function, \(Q^\pi(s,a)\), is defined as the expected reward of taking action
\(a\) in state \(s\) and then following policy \(\pi\) thereafter:
\(Q^\pi(s_0, a_0) \defeq \E_\pi \left[ \sum_{t=0}^\infty \gamma^t r_t \vert s_0=S, a_0=A \right]\).
The Bellman operator on \(Q^\pi\) is defined as
\(\B[Q^\pi](s_t,a_t) \defeq \E_{P(s_{t+1} \vert s_t, a_t)\pi(a_{t+1} \vert s_t)}
\left[ r(s_t, a) + \gamma Q^\pi(s_{t+1}, a_{t+1}) \right]\). With the
\(Q\)-function we can define a policy that always picks the action with the 
highest \(Q\)-value in any state. In deep RL, we model \(Q\) with a
neural network \(\hat{Q}_\omega\). One way to facilitate exploration in this
policy is to introduce uncertainty in the
\(Q\)-value function.

@@

We would like to approximate the posterior distribution of the Q-value function.
\citet{osband_why_2017} prove that in finite horizon MDPs posterior sampling for
reinforcement learning has at least a near-optimal regret bound. They further conjecture
that this cound can be further improved to show an optimal regret bound. Additionally,
a posterior sampling based reinfrocement learning algorithm could be made to utilize
prior information or domain knowledge through an appropriate prior. This should improve
policy convergence rate.
Combined with the computational efficiency of posterior sampling, this motivates the
development of a Bayesian reinforcement learning algorithm with functional priors.

We can use the functional variational bayesian neural
network (FVBNN)~\citep{sun_functional_2019} framework discussed earlier.
NoisyNet uses one sample from Q for each optimization step. We will instead use
\(k\) samples from Q. This lets us add the KL-term from FVBNN to the loss function that
encourages stochasticity.
The maximization target becomes:
\begin{equation}
    \log p(\D_\omega^N \vert f) - \KL \left[ q(\f^{\D_\omega^N}, \f^\mathbf{M}) \Vert p(\f^{\D_\omega^N}, \f^\mathbf{M}) \right],
\end{equation}
where \(\f^{\D_\omega^N}\) is \(f\) applied to the dataset \(\D_\omega^N\), and \(\f^\mathbf{M}\) is \(f\) applied to
a set of i.i.d. random points \(\mathbf{M} = \{m_i \sim c \mid i=1,\dots,k\} \subseteq \state \times \action\)
with full support. We will let the measurement \(\mathbf{M}\) be the the set of all state-action
pairs for states we have already explored. If the MDP is ergodic, then \(\textbf{M}\) will eventually
have full support in \(\state \times \action\).

@@

\begin{algorithm}
\caption{Functional Bayesian RL}
\begin{algorithmic}
    \State \(\D \gets \emptyset\)
    \State \(s \sim P_0\)
    \State Initialize \(\phi, \omega\)
    \While{not converged}
        \State \(f \sim q_\phi\)
        \State \(a \sim \arg\max_a f(s,a)\)
        \State \(s' \sim P(\cdot \vert s,a)\)
        \State \(r = r(s',a,s)\)
        \State \(\D \gets \D \cup \{s,a,r,s'\}\)
        \State \Call{UpdatePosterior}{$\phi, \omega, \D$}
    \EndWhile
    \Statex
    \Function{UpdatePosterior}{$\phi, \omega, \D$}
        \State \(T \sim \D\)
        \State \(\f \gets \{f_i \sim q_\phi \mid i=1,\dots,k\}\)
        \State \(\mathbf{b} \gets \{b_i \sim q_\omega \mid i=1,\dots,k\}\)
        \State \(\mathcal{F} \gets \emptyset\)
        \State \(\Delta_\mathcal{L} \gets 0\)
        \State \(S \gets 0\)
        \For{\(\{s,a,r,s'\} \in T\)}
            \State \(\bm{q_\mu}, \bm{q_\sigma} \gets \max_a \f(s,a)\)
            \State \(\mathcal{F} \gets \mathcal{F} \cup \f(s,a)\)
            \State \(\bm{G} \gets \{r + \gamma b_i(s,a) \mid i = 1, 
        \dots, k\}\)
            \State \(S \gets S + \frac{1}{k}p(\bm{q} \vert s, a)\)
            \State \(\Delta_\mathcal{L} \gets \Delta_\mathcal{L} - \frac{1}{\lvert k \rvert} \nabla_\phi \log p(\bm{q_\mu} \vert \bm{G}, \bm{q_\sigma})\)
        \EndFor
        \State \(\Delta_\KL \gets S - SSGE(\mathcal{F})\)
        \State \(\phi \gets \phi + \eta_\phi (\Delta_\mathcal{L} - \Delta_\KL)\)
        \State \(\omega \gets \omega + \eta_\omega (\omega - \phi)\)
    \EndFunction
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/Cartpole_score}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/MountainCar_score}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/Cartpole_score}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/MountainCar_score}
    \end{subfigure}
    \caption{Example figure 1}
    \label{fig:fig_1}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{example-image}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{example-image}
    \end{subfigure}
    \includegraphics[width=\columnwidth]{example-image}
    \caption{Performance on the CartPole-v1 environment.}
\end{figure}
